### 2011-05-13 paper/paper.tex
Minor edits + conclusion edit

We consider the problem of randomly and uniformly sampling a point from a convex region in $n$-dimensional space. This has applications, for instance, in convex optimization \cite{Dabbene}. Sampling can also allow us to estimate the volume of a convex region \cite{Dyer}.

We phrase the problem as follows: we are given a convex region $\mathcal{R}$ in the form of a \emph{membership oracle} which determines whether a given point $x$ is in $\mathcal{R}$, along with a starting point $x_0$ in $\mathcal{R}$. The task is to sample a point from the region with with a distribution as close to uniform as possible.

@@

The hit-and-run method is another random walk, but it has the advantage that it is guaranteed (almost surely) to move to another point. At a point $x$, we first choose a random direction, uniformly. Now consider the two-sided line $\ell$ which goes through $x$ which goes in this direction. Choose a point uniformly randomly from $\ell\cap\mathcal{R}$, and jump to that point.

@@

The Metropolis methods both have the same basic structure. A transition in the Markov chain works simply by choosing another point from some distribution centered at our current location; if that point is in the convex region $\mathcal{R}$, then we move to that point; otherwise, we stay. The only difference between the two methods is the distribution we use. Ball-walking chooses a point randomly from a ball, and Gaussian-walking chooses each coordinate from a Gaussian.

@@

\begin{theorem}
Any Markov chain whose state space is an open subset of $\mathbb{R}^n$ with a symmetric transition distribution has the uniform distribution as a stationary distribution.
\label{symmetric}
\end{theorem}
\begin{proof}
We will let $S$ be the set of all states of the chain. Furthermore, let $\mu$ be the uniform distribution on $S$, i.e. $\mu(x) = \frac{1}{\text{Vol}(S)}$, where $\text{Vol}(S)$ denotes the volume of $S$. Then we want to show that for every subset $A \subseteq S$, $\int_A (\mu P)(y)\, dy = \frac{\text{Vol}(A)}{\text{Vol}(S)}$. To do this, notice the following:
\begin{align*}
\int_A (\mu P)(y)\, dy &= \int_A \int_S \mu(x)P(x,y)\, dx\, dy \\
&= \int_S \int_A \mu(y)P(y,x)\, dy\, dx
\end{align*}
where here we have used the fact that $\mu(x) = \mu(y)$, $P(x,y) = P(y,x)$, and that we can interchange the order of integration by Fubini's theorem because the double integral is a probability, and therefore finite and the integrand is always positive. However, we can see that this last quantity is the probability of starting at a point in $A$ and moving to a point in $S$. Since we will always move to a point in $S$, that quantity is just the probability that we start in $A$, which is simply $\int_A \mu(y)\, dy = \frac{\text{Vol}(A)}{\text{Vol}(S)}$, as desired.
\end{proof}
\begin{remark} 
\emph{This is a continuous analog of the following fact: given an undirected $d$-regular graph $G$, consider a Markov chain which, from any vertex, jumps to each of its neighbors with probability $1/d$. Then the stationary distribution on this Markov chain is the stationary distribution.}
\end{remark}

@@

We now consider the hit-and-run method for sampling. Recall that the idea behind this method is to transition as follows: from a point $x \in \mathcal{R}$, randomly choose a line $\ell$ through $x$, and then randomly choose a point on $\ell \cap \mathcal{R}$ to jump to.

@@

One potential drawback of this method is that it needs to find $\ell \cap \mathcal{R}$ given $\ell$. Given a membership oracle for $\mathcal{R}$, we can find the endpoints (up to some specified precision) of $\ell$ with a binary search. Any step of the hit-and-run method takes much longer than any step of the ball-walking method. Thus, the hit-and-run method is only an improvement in the number of transition steps that we want to use; if this is large enough to overcome the slowdown from the binary search (and in theory, it should be), then it will also have an improved runtime.

@@

In general, the adaptive methods performed better than their non-adaptive counterparts. When started from the corner of a cube or a simplex, the non-adaptive methods would always stay in that corner for high-dimension test cases, even $n=10$. The adaptive methods, however, produced a distribution which looked much more uniform. However, in cases where we do not start close to a boundary, the adaptive methods do not have as much of an advantage over the non-adaptive methods.

@@

\bibitem{Dyer} Dyer, M.E., Frieze, A.M., and Kannan, R. ``A random polynomial-time algorithm for approximating the volume of convex bodies." Journal of the ACM, 38:1-17, 1991.

@@

\bibitem{Vempala} Lov\'asz, L. and Vempala, S. ``Hit-and-Run from a corner." \emph{SIAM Journal on Computing}, 35(4):9851005, 2006.

@@

\bibitem{Smith} Smith, R.L. ``Efficient Monte-Carlo procedures for generating points uniformly distributed over
bounded regions." Oper. Res., 32 (1984), pp. 1296-1308.

>>>>

We consider the problem of randomly and uniformly sampling a point from a convex region in $n$-dimensional space. This has applications, for instance, in convex optimization \cite{Dabbene}. We phrase the problem as follows: we are given a convex region $\mathcal{R}$ in the form of a \emph{membership oracle} which determines whether a given point $x$ is in $\mathcal{R}$, along with a starting point $x_0$ in $\mathcal{R}$. The task is to sample a point from the region with with a distribution as close to uniform as possible.

@@

The hit-and-run method is another random walk, but it has the advantage that it is guaranteed (almost surely) to move to another point. At a point $x$, we first choose a random direction, uniformly. Now consider the two-sided line $\ell$ which goes through $x$ and in the chosen direction. Choose a point uniformly randomly from $\ell\cap\mathcal{R}$, and jump to that point.

@@

The Metropolis methods both have the same basic structure. A transition in the Markov chain works simply by choosing another point from some distribution centered at our current location; if that point is in the convex region $\mathcal{R}$, then we move to that point; otherwise, we stay. The only difference between the two methods is the distribution we use. Ball-walking chooses a point uniformly from a ball, and Gaussian-walking chooses each coordinate from a Gaussian.

@@

\begin{theorem}
Any Markov chain whose state space is an open subset of $\mathbb{R}^n$ with a symmetric transition distribution has the uniform distribution as a stationary distribution.
\label{symmetric}
\end{theorem}
\begin{proof}
We will let $S$ be the set of all states of the chain. Furthermore, let $\mu$ be the uniform distribution on $S$, i.e. $\mu(x) = \frac{1}{\text{Vol}(S)}$, where $\text{Vol}(S)$ denotes the volume of $S$. Then we want to show that for every subset $A \subseteq S$, $\int_A (\mu P)(y)\, dy = \frac{\text{Vol}(A)}{\text{Vol}(S)}$. To do this, notice the following:
\begin{align*}
\int_A (\mu P)(y)\, dy &= \int_A \int_S \mu(x)P(x,y)\, dx\, dy \\
&= \int_S \int_A \mu(y)P(y,x)\, dy\, dx
\end{align*}
where here we have used the fact that $\mu(x) = \mu(y)$, $P(x,y) = P(y,x)$, and that we can interchange the order of integration by Fubini's theorem because the double integral is a probability, and therefore finite, and the integrand is always positive. However, we can see that this last quantity is the probability of starting at a point in $A$ and moving to a point in $S$. Since we will always move to a point in $S$, that quantity is just the probability that we start in $A$, which is simply $\int_A \mu(y)\, dy = \frac{\text{Vol}(A)}{\text{Vol}(S)}$, as desired.
\end{proof}
\begin{remark} 
\emph{This is a continuous analog of the following fact: given an undirected $d$-regular graph $G$, consider a Markov chain which, from any vertex, jumps to each of its neighbors with probability $1/d$. Then the stationary distribution on this Markov chain is the stationary distribution.}
\end{remark}

@@

We now consider the hit-and-run method for sampling. Recall that the idea behind this method is to transition as follows: from a point $x \in \mathcal{R}$, uniformly choose a line $\ell$ through $x$, and then uniformly choose a point on $\ell \cap \mathcal{R}$ to jump to.

@@

One potential drawback of this method is that it needs to find $\ell \cap \mathcal{R}$ given $\ell$. Given a membership oracle for $\mathcal{R}$, we can find the endpoints (up to some specified precision) of $\ell$ with a binary search. Any step of the hit-and-run method takes much longer than any step of the ball-walking method. Thus, the hit-and-run method is only an improvement in the number of transition steps that we want to use. if this is large enough to overcome the slowdown from the binary search (and in theory, it should be), then it will also have an improved runtime.

@@

We found that the adaptive methods did not give a significant boost to the performance of Metropolis methods. When started in a bad location, the adaptive methods often shrank their step size down to a point where it was no longer likely to spread to all portions of the region. However, when starting in the middle of a large region, the adaptive methods performed better than their nonadaptive counterparts, as they were able to spread out much more rapidly. In practice, adaptive methods may prove useful when the algorithm has no sense of what size the region should be, but a suitably chosen step size will usually perform better if a very rough size of the object is known.
%In general, the adaptive methods performed better than their non-adaptive counterparts. When started from the corner of a cube or a simplex, the non-adaptive methods would always stay in that corner for high-dimension test cases, even $n=10$. The adaptive methods, however, produced a distribution which looked much more uniform. However, in cases where we do not start close to a boundary, the adaptive methods do not have as much of an advantage over the non-adaptive methods.

@@

@@

\bibitem{Vempala} Lov\'asz, L. and Vempala, S. Hit-and-Run from a corner. \emph{SIAM Journal on Computing}, 35(4):9851005, 2006.

@@

\bibitem{Smith} Smith, R.L. \emph{Efficient Monte-Carlo procedures for generating points uniformly distributed over
bounded regions}, Oper. Res., 32 (1984), pp. 1296-1308.
